### 2018-01-17 circuit-elements/paper.tex
abstract + 1st question left

	The resistance of the thermistor at various temperatures was measured and recorded. The uncertainty, as specified by the manufacturer, was $ \pm (0.2\% + 1) $ for resistances in the $ (2 \si{\kilo\ohm}, 200 \si{\kilo\ohm}) $ range. The uncertainty in temperature measurements was taken to be $ \pm 0.5 \si{\kelvin} $. \vspace{1em}

	\paperfig{thermistor}{\begin{papertable}{|[outer]I|[inner]C|[inner]U|[outer]}
			\paperoline\papertableindexheader&\textsc{Temperature}&\multicolumn{2}{c|[outer]}{\textsc{Resistance}}\\
			\papertablecheadersymbol{T$ $(\pm0.5$ $\si{\kelvin})}\papertableuheadersymbole{R$ $(\si{\kilo\ohm})}\\\paperiline
			\papertableindex\papertablecval{352.5}\papertableuval{6.68}{.02}\\\paperiline
			\papertableindex\papertablecval{349.0}\papertableuval{6.93}{.02}\\\paperiline
			\papertableindex\papertablecval{348.0}\papertableuval{7.54}{.03}\\\paperiline
			\papertableindex\papertablecval{346.0}\papertableuval{8.01}{.03}\\\paperiline
			\papertableindex\papertablecval{340.0}\papertableuval{9.95}{.03}\\\paperiline
			\papertableindex\papertablecval{338.0}\papertableuval{10.33}{.03}\\\paperiline
			\papertableindex\papertablecval{335.0}\papertableuval{11.76}{.03}\\\paperiline
			\papertableindex\papertablecval{334.5}\papertableuval{12.36}{.03}\\\paperiline
			\papertableindex\papertablecval{329.5}\papertableuval{14.45}{.04}\\\paperiline
			\papertableindex\papertablecval{327.5}\papertableuval{15.33}{.04}\\\paperiline
			\papertableindex\papertablecval{324.0}\papertableuval{17.23}{.04}\\\paperiline
			\papertableindex\papertablecval{321.0}\papertableuval{19.51}{.05}\\\paperiline
			\papertableindex\papertablecval{319.0}\papertableuval{21.7}{.1}\\\paperiline
			\papertableindex\papertablecval{317.0}\papertableuval{24.3}{.1}\\\paperiline
			\papertableindex\papertablecval{314.0}\papertableuval{26.1}{.2}\\\paperiline
			\papertableindex\papertablecval{309.5}\papertableuval{31.4}{.2}\\\paperiline
			\papertableindex\papertablecval{308.5}\papertableuval{33.7}{.2}\\\paperiline
			\papertableindex\papertablecval{307.5}\papertableuval{35.5}{.2}\\\paperiline
			\papertableindex\papertablecval{305.0}\papertableuval{38.0}{.2}\\\paperiline
			\papertableindex\papertablecval{304.0}\papertableuval{40.3}{.2}\\\paperiline
			\papertableindex\papertablecval{301.5}\papertableuval{44.3}{.2}\\\paperiline
			\papertableindex\papertablecval{299.5}\papertableuval{47.7}{.2}\\\paperiline
			\papertableindex\papertablecval{298.0}\papertableuval{51.6}{.2}\\\paperiline
			\papertableindex\papertablecval{296.0}\papertableuval{56.3}{.2}\\\paperiline
			\papertableindex\papertablecval{294.0}\papertableuval{61.3}{.2}\\\paperiline
			\papertableindex\papertablecval{292.5}\papertableuval{66.9}{.2}\\\paperiline
			\papertableindex\papertablecval{290.5}\papertableuval{73.8}{.2}\\\paperiline
			\papertableindex\papertablecval{286.0}\papertableuval{92.9}{.3}\\\paperiline
			\papertableindex\papertablecval{283.0}\papertableuval{104.7}{.3}\\\paperiline
			\papertableindex\papertablecval{278.0}\papertableuval{~126.2}{\parbox{\widthof{126.2~}}{.4}}\\\paperoline
		\end{papertable}\vspace{-1.5em}}
	{Resistance of thermistor at the given temperatures. As expected, resistance decreased with increasing temperature.}\vspace{1em}

	Further, current and potential for a tungsten filament light bulb were measured and recorded. As specified by the manufacturer, the uncertainty in potential measurements was taken to be \( \pm (0.25\% + 1) \); current measurements carried an uncertainty of \( \pm (1\% + 3) \). \vspace{1em}

	\paperfig{light-bulb}{\begin{papertable}{|[outer]I|[inner]U|[inner]U|[outer]}
			\paperoline\papertableindexheader&\multicolumn{2}{c|[inner]}{\textsc{Potential}}&\multicolumn{2}{c|[outer]}{\textsc{Current}}\\
			\papertableuheadersymbol{V$ $(\si{\volt})}\papertableuheadersymbole{I$ $(\si{\milli\ampere)}}\\\paperiline
			\papertableindex\papertableuval{0.0222}{.0002}\papertableuval{0.324}{.003}\\\paperiline
			\papertableindex\papertableuval{0.0776}{.0003}\papertableuval{1.045}{.009}\\\paperiline
			\papertableindex\papertableuval{0.1122}{.0004}\papertableuval{1.39}{.01}\\\paperiline
			\papertableindex\papertableuval{0.244}{.002}\papertableuval{2.27}{.03}\\\paperiline
			\papertableindex\papertableuval{0.472}{.002}\papertableuval{3.40}{.04}\\\paperiline
			\papertableindex\papertableuval{0.541}{.002}\papertableuval{3.69}{.04}\\\paperiline
			\papertableindex\papertableuval{0.697}{.003}\papertableuval{4.28}{.04}\\\paperiline
			\papertableindex\papertableuval{0.871}{.003}\papertableuval{4.89}{.05}\\\paperiline
			\papertableindex\papertableuval{1.128}{.004}\papertableuval{5.72}{.05}\\\paperiline
			\papertableindex\papertableuval{1.840}{.006}\papertableuval{7.68}{.07}\\\paperiline
			\papertableindex\papertableuval{1.940}{.006}\papertableuval{7.93}{.07}\\\paperiline
			\papertableindex\papertableuval{4.01}{.02}\papertableuval{12.3}{.1}\\\paperiline
			\papertableindex\papertableuval{6.00}{.03}\papertableuval{15.8}{.1}\\\paperiline
			\papertableindex\papertableuval{8.01}{.03}\papertableuval{18.8}{.2}\\\paperiline
			\papertableindex\papertableuval{10.24}{.04}\papertableuval{21.8}{.3}\\\paperiline
			\papertableindex\papertableuval{12.07}{.04}\papertableuval{24.1}{.3}\\\paperiline
			\papertableindex\papertableuval{14.27}{.05}\papertableuval{26.6}{.3}\\\paperiline
			\papertableindex\papertableuval{16.01}{.05}\papertableuval{28.5}{.3}\\\paperiline
			\papertableindex\papertableuval{17.98}{.05}\papertableuval{30.6}{.3}\\\paperiline
			\papertableindex\papertableuval{18.45}{.06}\papertableuval{31.1}{.3}\\\paperiline
			\papertableindex\papertableuval{19.86}{.06}\papertableuval{32.5}{.3}\\\paperiline
			\papertableindex\papertableuval{21.0}{.2}\papertableuval{33.6}{.4}\\\paperiline
			\papertableindex\papertableuval{22.1}{.2}\papertableuval{34.6}{.4}\\\paperiline
			\papertableindex\papertableuval{22.9}{.2}\papertableuval{35.4}{.4}\\\paperiline
			\papertableindex\papertableuval{24.0}{.2}\papertableuval{36.4}{.4}\\\paperiline
			\papertableindex\papertableuval{25.1}{.2}\papertableuval{37.3}{.4}\\\paperoline
			\end{papertable}\vspace{-1.5em}}
	{Potential and current through the incandescent bulb. Evidently, increased current through the filament led to an increase in potential.} \vspace{1em}

@@

	\paperfig{thermistor-plot}{\pdf{Thermistor}}{Resistance of the thermistor plotted against the temperature in the thermal flask, with least squares fit included in blue. Error bars on both temperature and resistance were too small to be visible. The plot of residuals indicated that the model is a better fit for greater temperatures.} \vspace{1em}

	A visualization of residuals clearly indicated that the data can be better modelled for greater temperatures. Indeed, least squares fitting performed without the data points corresponding to three lowest temperatures yielded a reduced \( \chi ^ 2 \) value of \( 181.5543 \). % Removing the six lower-temperature points led to a model with reduced \( \chi ^ 2 \) value of \( 130.1026 \).

@@

	Upon obtaining the value of \( a \), the universal constant \( n \) in Stefan-Boltzmmann law could be approximated. Plugging \( a = 1.63 \pm 0.01 \) into (6) yielded \( 4.17 \pm 0.05 \), which was close the actual value of \( n = 4\). Another constant appearing in Stefan-Boltzmann law, \( \epsilon \) (emissivity of the radiating body could be determined), differs depending on the material. \( \epsilon = 1 \) corresponds to a body that is a perfect absorber and emitter of radiant energy (i.e., a \textit{black body}); \( \epsilon = 0 \) corresponds to a perfect reflector of radiant energy.

@@

	Conclusion

@@

	Sources

>>>>

	The resistance of the thermistor at various temperatures was measured and recorded. The uncertainty, as specified by the manufacturer, was $ \pm (0.2\% + 1 $ LSG) for resistances in the $ (2 \si{\kilo\ohm}, 200 \si{\kilo\ohm}) $ range. The uncertainty in temperature measurements was taken to be $ \pm 0.5 \si{\kelvin} $. \vspace{1em}

	\paperfig{thermistor}{\begin{papertable}{|[outer]I|[inner]C|[inner]U|[outer]}
			\paperoline\papertableindexheader&\textsc{Temperature}&\multicolumn{2}{c|[outer]}{\textsc{Resistance}}\\
			\papertablecheadersymbol{T$ $(\pm0.5$ $\si{\kelvin})}\papertableuheadersymbole{R$ $(\si{\kilo\ohm})}\\\paperiline
			\papertableindex\papertablecval{352.5}\papertableuval{6.68}{.02}\\\paperiline
			\papertableindex\papertablecval{349.0}\papertableuval{6.93}{.02}\\\paperiline
			\papertableindex\papertablecval{348.0}\papertableuval{7.54}{.03}\\\paperiline
			\papertableindex\papertablecval{346.0}\papertableuval{8.01}{.03}\\\paperiline
			\papertableindex\papertablecval{340.0}\papertableuval{9.95}{.03}\\\paperiline
			\papertableindex\papertablecval{338.0}\papertableuval{10.33}{.03}\\\paperiline
			\papertableindex\papertablecval{335.0}\papertableuval{11.76}{.03}\\\paperiline
			\papertableindex\papertablecval{334.5}\papertableuval{12.36}{.03}\\\paperiline
			\papertableindex\papertablecval{329.5}\papertableuval{14.45}{.04}\\\paperiline
			\papertableindex\papertablecval{327.5}\papertableuval{15.33}{.04}\\\paperiline
			\papertableindex\papertablecval{324.0}\papertableuval{17.23}{.04}\\\paperiline
			\papertableindex\papertablecval{321.0}\papertableuval{19.51}{.05}\\\paperiline
			\papertableindex\papertablecval{319.0}\papertableuval{21.7}{.1}\\\paperiline
			\papertableindex\papertablecval{317.0}\papertableuval{24.3}{.1}\\\paperiline
			\papertableindex\papertablecval{314.0}\papertableuval{26.1}{.2}\\\paperiline
			\papertableindex\papertablecval{309.5}\papertableuval{31.4}{.2}\\\paperiline
			\papertableindex\papertablecval{308.5}\papertableuval{33.7}{.2}\\\paperiline
			\papertableindex\papertablecval{307.5}\papertableuval{35.5}{.2}\\\paperiline
			\papertableindex\papertablecval{305.0}\papertableuval{38.0}{.2}\\\paperiline
			\papertableindex\papertablecval{304.0}\papertableuval{40.3}{.2}\\\paperiline
			\papertableindex\papertablecval{301.5}\papertableuval{44.3}{.2}\\\paperiline
			\papertableindex\papertablecval{299.5}\papertableuval{47.7}{.2}\\\paperiline
			\papertableindex\papertablecval{298.0}\papertableuval{51.6}{.2}\\\paperiline
			\papertableindex\papertablecval{296.0}\papertableuval{56.3}{.2}\\\paperiline
			\papertableindex\papertablecval{294.0}\papertableuval{61.3}{.2}\\\paperiline
			\papertableindex\papertablecval{292.5}\papertableuval{66.9}{.2}\\\paperiline
			\papertableindex\papertablecval{290.5}\papertableuval{73.8}{.2}\\\paperiline
			\papertableindex\papertablecval{286.0}\papertableuval{92.9}{.3}\\\paperiline
			\papertableindex\papertablecval{283.0}\papertableuval{104.7}{.3}\\\paperiline
			\papertableindex\papertablecval{278.0}\papertableuval{~126.2}{\parbox{\widthof{126.2~}}{.4}}\\\paperoline
		\end{papertable}\vspace{-1.5em}}
	{Resistance of thermistor at the given temperatures. All uncertainties were propagated using the method of quadratures.}\vspace{1em}

	Further, current and potential for a tungsten filament light bulb were measured and recorded. As specified by the manufacturer, the uncertainty in potential measurements was taken to be \( \pm (0.25\% + 1 \) of last displayed significant digit); current measurements carried an uncertainty of \( \pm (1\% + 3 \) of last displayed significant digit). \vspace{1em}

	\paperfig{light-bulb}{\begin{papertable}{|[outer]I|[inner]U|[inner]U|[outer]}
			\paperoline\papertableindexheader&\multicolumn{2}{c|[inner]}{\textsc{Potential}}&\multicolumn{2}{c|[outer]}{\textsc{Current}}\\
			\papertableuheadersymbol{V$ $(\si{\volt})}\papertableuheadersymbole{I$ $(\si{\milli\ampere)}}\\\paperiline
			\papertableindex\papertableuval{0.0222}{.0002}\papertableuval{0.324}{.003}\\\paperiline
			\papertableindex\papertableuval{0.0776}{.0003}\papertableuval{1.045}{.009}\\\paperiline
			\papertableindex\papertableuval{0.1122}{.0004}\papertableuval{1.39}{.01}\\\paperiline
			\papertableindex\papertableuval{0.244}{.002}\papertableuval{2.27}{.03}\\\paperiline
			\papertableindex\papertableuval{0.472}{.002}\papertableuval{3.40}{.04}\\\paperiline
			\papertableindex\papertableuval{0.541}{.002}\papertableuval{3.69}{.04}\\\paperiline
			\papertableindex\papertableuval{0.697}{.003}\papertableuval{4.28}{.04}\\\paperiline
			\papertableindex\papertableuval{0.871}{.003}\papertableuval{4.89}{.05}\\\paperiline
			\papertableindex\papertableuval{1.128}{.004}\papertableuval{5.72}{.05}\\\paperiline
			\papertableindex\papertableuval{1.840}{.006}\papertableuval{7.68}{.07}\\\paperiline
			\papertableindex\papertableuval{1.940}{.006}\papertableuval{7.93}{.07}\\\paperiline
			\papertableindex\papertableuval{4.01}{.02}\papertableuval{12.3}{.1}\\\paperiline
			\papertableindex\papertableuval{6.00}{.03}\papertableuval{15.8}{.1}\\\paperiline
			\papertableindex\papertableuval{8.01}{.03}\papertableuval{18.8}{.2}\\\paperiline
			\papertableindex\papertableuval{10.24}{.04}\papertableuval{21.8}{.3}\\\paperiline
			\papertableindex\papertableuval{12.07}{.04}\papertableuval{24.1}{.3}\\\paperiline
			\papertableindex\papertableuval{14.27}{.05}\papertableuval{26.6}{.3}\\\paperiline
			\papertableindex\papertableuval{16.01}{.05}\papertableuval{28.5}{.3}\\\paperiline
			\papertableindex\papertableuval{17.98}{.05}\papertableuval{30.6}{.3}\\\paperiline
			\papertableindex\papertableuval{18.45}{.06}\papertableuval{31.1}{.3}\\\paperiline
			\papertableindex\papertableuval{19.86}{.06}\papertableuval{32.5}{.3}\\\paperiline
			\papertableindex\papertableuval{21.0}{.2}\papertableuval{33.6}{.4}\\\paperiline
			\papertableindex\papertableuval{22.1}{.2}\papertableuval{34.6}{.4}\\\paperiline
			\papertableindex\papertableuval{22.9}{.2}\papertableuval{35.4}{.4}\\\paperiline
			\papertableindex\papertableuval{24.0}{.2}\papertableuval{36.4}{.4}\\\paperiline
			\papertableindex\papertableuval{25.1}{.2}\papertableuval{37.3}{.4}\\\paperoline
			\end{papertable}\vspace{-1.5em}}
	{Potential and current through the incandescent bulb. All uncertainties were propagated using the method of quadratures.} \vspace{1em}

@@

	\paperfig{thermistor-plot}{\pdf{Thermistor}}{Resistance and residuals plotted against the temperature in the thermal flask, with least squares fit included in blue. Error bars on both temperature and resistance were too small to be visible.} \vspace{1em}

	A visualization of residuals clearly indicated that the data can be better modelled for greater temperatures. Indeed, least squares fitting performed without the data points corresponding to three lowest temperatures yielded a reduced \( \chi ^ 2 \) value of \( 181.5543 \). Removing more lower-temperature data points further improved reduced \( \chi ^ 2 \). % Removing the six lower-temperature points led to a model with reduced \( \chi ^ 2 \) value of \( 130.1026 \).

@@

	As evidenced above, thermistors and metal resistors have different variations of resistance with temperatures. For metal resistors, resistance increasing with temperature is intuitively explained TODO!

	Upon obtaining the value of \( a \), the universal constant \( n \) in Stefan-Boltzmmann law could be approximated. Plugging \( a = 1.63 \pm 0.01 \) into (6) yielded \( 4.17 \pm 0.05 \) (uncertainty propagated using the method  of quadratures), which was close the actual value of \( n = 4\). Another constant appearing in Stefan-Boltzmann law, \( \epsilon \) (emissivity of the radiating body), differs depending on the material. \( \epsilon = 1 \) corresponds to a body that is a perfect absorber and emitter of radiant energy (i.e., a \textit{black body}); \( \epsilon = 0 \) corresponds to a perfect reflector of radiant energy.

@@

		This experiment has investigated two nonlinear circuit elements, a thermistor and an incandescent tungsten filament light bulb. Resistance of the thermistor was found to be a function of the temperature, satisfying the exponential relation in (1) with \( R_0 = (1.0 \pm 0.1) \times 10^{13} \si{\ohm} \) and \( T_0 = (24.7 \pm 0.3) \si{\kelvin} \).  For the light bulb, the relationship between potential and current was shown to follow a power law introduced in (2), with \( k =  (0.067 \pm 0.002) \si{\volt\per\ampere} \) and \( a = 1.63 \pm 0.01 \). The most significant sources of error arose from potential uneven temperature of water in the thermal flask with the thermistor and inconsistencies in current and potential readings of the multimeter when switching the range of input signal.

		The above value of \( a \) was further used to approximate the power \( n \) of \( T \) in Stefan-Boltzmann law; \( n \) was found to be \( 4.17 \pm 0.05 \), reasonably close to the actual value of \( n = 4 \).

@@

	\papersource{Serbanescu, R., Non-linear Circuit Elements, 2015}
